\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-01-18 eg6-letters/NIM/EG6_RTPC_NIM.tex
Editing...

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A New Radial Time 
Projection Chamber for CLAS at JLab}}
\input author_list.tex

@@

\begin{abstract}
A new Radial Time Projection Chamber (RTPC) was developed at Jefferson 
laboratory (JLab) to track low-energy nuclear recoils for the purpose of 
measuring nuclear exclusive channels, such as, coherent Deeply Virtual Compton 
Scattering (DVCS) and coherent production of mesons on $^4$He. In such 
processes, $^4$He nucleus remains intact and recoils as a whole in the final 
state. The CEBAF Large Acceptance Spectrometer (CLAS) at JLab has been used for 
similar studies on proton targets, however it cannot track recoils with momenta 
less than 250 MeV/c. In 2009, we carried out measurements on $^4$He target 
using CLAS spectrometer, which was upgraded with the RTPC during the CLAS-EG6 
experimental run. The detector was positioned directly around a gaseous target, 
allowing a detection threshold of $~$100  MeV/c for $^4$He. This article 
discuses the RTPC design, work principle, developed calibration methods and 
detector performance.
\end{abstract}

@@

The CEBAF Large Acceptance Spectrometer (CLAS)~\cite{CLASref} is installed in 
the Hall-B of the Thomas Jefferson National Accelerator Facility (TJNAF), 
Newport News, Virginia, USA. The accelerator of TJNAF provides, simultaneously, 
a high power electron beam of energy up to 6 GeV and 100$\%$ duty cycle to 
three experimental Halls (A, B, C). The CLAS spectrometer, located in Hall-B, 
is composed of several sub-detectors and two magnets. The spectrometer was 
designed to track charged particles of momentum greater than 250 MeV/c as well 
as detecting real photons. Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the CLAS detector, ordered with respect to their radial 
distance from the beam line as:

\begin{itemize}
 \item Three regions of Drift Chambers (DC) for tracking of charged 
    particles~\cite{DCref}.
 \item Superconducting toroidal magnet which enables the tracking in the DC by 
    bending the trajectories.
 \item Cherenvkov Counters (CC) to help in separating electrons from negative 
    particles, such as pions ~\cite{CCref}.
 \item Scintillation Counters (SC) to identify hadrons via measuring their 
    Time-Of-Flight (TOF)~\cite{TOFref}.
 \item Forward Electromagnetic Calorimeter (EC) which measures the energy of 
    the showering particles (electrons and photons) and neutrons~\cite{ECref}.
\end{itemize}

\begin{figure}[tbp]
\centering \includegraphics[scale=0.3]{fig/test_clas.png}
\caption{A three dimensional representation of the primary CLAS setup 
   represents the geometry and the relative position of each subsystem. From 
   outside, the electromagnetic calorimeter (green), time-of-flight system 
   (red), Cherenkov counters (purple), three drift chamber regions (blue) and 
the superconducting torus coils (yellow).} \label{fig:CLAS}
\end{figure}

The DVCS process is the exclusive electroproduction of a real photon from a 
quark of a hadronic target ($H$), $eH \rightarrow e' H' \gamma$. In the basic 
setup of CLAS, the real photons are detected by the forward EC, which covers 
polar angles from 8$^{\circ}$ to 45$^{\circ}$. With a 6 GeV electron beam, the 
majority of the DVCS photons are produced at polar angles below 15$^{\circ}$, 
where the acceptance of the EC is poor. In CLAS-E1DVCS experiment (2005), CLAS 
was upgraded with the addition of an Inner Calorimeter (IC). The IC is 
constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long ,
corresponding to 17 radiation lengths, with a 1.33~$\times$~1.33 cm$^2$ front 
surface and a 1.6~$\times$~1.6 cm$^2$ back surface. The achieved energy 
resolution is around 3 to 4$\%$ for photon energies between 2 GeV and 5 GeV and 
the angular resolution is between 3 to 5 mrad for the same energy range 
\cite{Hyon-suk}.

>>>>

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time 
Projection Chamber for $\alpha$ detection in CLAS at JLab}}
\input author_list.tex

@@

\begin{abstract}
A new Radial Time Projection Chamber (RTPC) was developed at Jefferson 
laboratory (JLab) to track low-energy nuclear recoils for the purpose of 
measuring nuclear exclusive channels, such as, coherent Deeply Virtual Compton 
Scattering (DVCS) and coherent production of mesons on $^4$He. In such 
processes, $^4$He nucleus remains intact and recoils as a whole in the final 
state. The CEBAF Large Acceptance Spectrometer (CLAS) at JLab has been used for 
similar studies on proton targets, however it cannot track recoils with momenta 
less than 250 MeV/c. In 2009, we carried out measurements on $^4$He target 
using CLAS spectrometer, which was upgraded with the RTPC during the CLAS-EG6 
experimental run. The detector was positioned directly around a gaseous target, 
allowing a detection threshold of $~$100  MeV/c for $^4$He. This article 
discuses the RTPC design, work principle, calibration methods and 
performances.
\end{abstract}

@@

The CEBAF Large Acceptance Spectrometer (CLAS)~\cite{CLASref} is installed in 
the Hall-B of the Thomas Jefferson National Accelerator Facility (TJNAF), 
Newport News, Virginia, USA. The accelerator of TJNAF provides, simultaneously, 
a high power electron beam of energy up to 6 GeV and 100$\%$ duty cycle to 
three experimental Halls (A, B, C). The CLAS spectrometer, located in Hall-B, 
is composed of several sub-detectors and two magnets. The spectrometer was 
designed to track charged particles of momentum greater than 250 MeV/c as well 
as detecting real photons. Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer, here is a list of the main elements
depicted in the figure:
\begin{itemize}
 \item The three regions of Drift Chambers (DC) used for the tracking of charged 
    particles~\cite{DCref} (in blue on the figure).
 \item The superconducting toroidal magnet (in yellow) bends the trajectories 
    of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item The Cherenvkov Counters (CC) used to separate electrons from other negative 
    particles, such as pions, at lower energies~\cite{CCref} (in pink).
 \item The Scintillation Counters (SC) used to identify hadrons by measuring their 
    Time-Of-Flight (TOF)~\cite{TOFref} (in red).
 \item The Electromagnetic Calorimeter (EC) used to separate electrons from other negative 
    particles at higher energies and to detect neutral particles (neutrons and 
    photons in particular)~\cite{ECref} (in green).
\end{itemize}

\begin{figure}[tbp]
\centering \includegraphics[scale=0.3]{fig/test_clas.png}
\caption{A three dimensional representation of the baseline CLAS setup. The
   full description is given in the text.} \label{fig:CLAS}
\end{figure}

The measurement of the DVCS process ($eH \rightarrow e' H' \gamma$, where 
$H$ is a hadron) necessitates some adaptations in this setup. Indeed, the EC 
covers only polar angles from 8$^{\circ}$ to 45$^{\circ}$ with a very small
acceptance at the lowest angles. With a 6 GeV electron beam on a proton target, the 
majority of the DVCS photons are produced at polar angles below 15$^{\circ}$, 
where the acceptance of the EC is poor. To resolve this issue an additional 
Inner Calorimeter (IC) has been built for the E01-113 experiment (2005). The IC is 
constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long,
corresponding to 17 radiation lengths. The achieved energy 
resolution is around 3 to 4$\%$ for photon energies between 2 GeV and 5 GeV and 
the angular resolution is between 3 to 5 mrad for the same energy range 
\cite{Hyon-suk}. The amount of background from Moller electron is however
very problematic at such small angles, so a 5 Tesla solenoid magnet has been
added around the target as a magnetic shield against the low energy background.
